---
title: "Project proposal"
author: "Team name"
output: html_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readxl)
```

## 1. Introduction

Community Concepts, INC. (CCI) is an non-profit organization based in Lewiston, Maine. Their focus is on providing resources, both information and financial, to local residents and businesses. Specifically, administering housing support, mental health support, family and children's services.

Data was provided by CCI in the form of two separate .xlsx files, one being Town Campaign Data and the other being Sept KPIs - Total Agency. This data details where CCI distributed their financial and informational help throughout the 2024 Fiscal Year. 

## 2. Data

Text goes here.

```{r}
sept_kpis <- read_excel("../data/FY24 Sept KPIs - Total Agency.xlsx")
town_campaign <- read_excel("../data/FY24 Town Campaign Data.xlsx")
```

## 3. Data analysis plan

Text goes here.


```{r}
# Code goes here
```


## 4. Data Ethics Review

The data ethics review section will be introduced in a separate class and is not part of the original proposal deadline.

